### 2024-11-22 paper/paper.qmd
initial figures

---
title: "Inflation of Egg Prices Across Canadian Grocery Stores"
author: Tara Chakkithara
date: today
format: pdf
warning: false
echo: false
---

```{r}
library(DBI)
library(RSQLite)
library(tidyverse)
library(lubridate)

connection <- dbConnect(RSQLite::SQLite(), "../data/data.sqlite")
script <- readLines("../scripts/egg_data.sql")
script <- paste(script, collapse = "\n")
data <- dbGetQuery(connection, script)
data$nowtime <- ymd_hms(data$nowtime)
data$current_price <- as.numeric(data$current_price)
```

## Data

The data for this analysis was sourced from Jacob Filipp's Project Hammer. 
### Vendor

```{r}
vendor_data <- data |>
  filter(!is.na(current_price)) |>
  mutate(month = format(nowtime, "%m")) |>
  select(vendor, current_price, month) |>
  group_by(vendor, month) |>
  summarise(avg_price = mean(current_price), .groups = "drop")

vendor_data |>
  ggplot(aes(x = month, y = avg_price, color = vendor, group = vendor)) +
  geom_point() +
  geom_line() +
  labs(x = "Month", y = "Average Egg Carton Price", color = "Vendor") +
  theme_minimal()
```

### Brand

```{r}
brand_data <- data |>
  filter(!is.na(current_price)) |>
  mutate(month = format(nowtime, "%m")) |>
  select(brand, current_price, month) |>
  group_by(brand, month) |>
  summarise(avg_price = mean(current_price), .groups = "drop")

brand_data |>
  ggplot(aes(x = month, y = avg_price, color = brand, group = brand)) +
  geom_point() +
  geom_line() +
  labs(x = "Month", y = "Average Egg Carton Price", color = "Brand") +
  theme_minimal()
```

### Egg Size

```{r}
egg_size_data <- data |>
  mutate(size = case_when(
    product_name == "Extra Large Size Free-Range Brown Eggs" ~ 'Extra Large',
    product_name == "Grade A Brown Eggs, Extra Large" ~ "Extra Large",
    product_name == "Grade A Eggs, Super Extra-Large" ~ "Extra Large",
    product_name == "Extra Large Free Run Brown Eggs" ~ "Extra Large",
    product_name == "Super Bon-Ee Grade A White Eggs, Super Extra Large" ~ "Super Extra Large",
    product_name == "Medium Grade A White Eggs" ~ "Small",
    product_name == "Naturegg Free Run Omega 3 Eggs" ~ "Small",
    product_name == "Omega 3 White Eggs" ~ "Small",
    product_name == "Super Bon-Ee Grade A Double Yolk White Eggs" ~ "Small",
    product_name == "Medium Size Free-Range Brown Eggs " ~ "Medium",
    TRUE ~ "Large"
  )) |>
  filter(!is.na(current_price)) |>
  mutate(month = format(nowtime, "%m")) |>
  select(size, current_price, month) |>
  group_by(size, month) |>
  summarise(avg_price = mean(current_price), .groups = "drop")

egg_size_data |>
  ggplot(aes(x = month, y = avg_price, color = size, group = size)) +
  geom_point() +
  geom_line() +
  labs(x = "Month", y = "Average Egg Carton Price", color = "Size") +
  theme_minimal()
```